Distributed Behavior in a Fluid Analogy Architecture

### 2017-12-02 papers/draft.tex
Enumerates a few of the pre-discussion points
Moves notes into paper structure

\begin{abstract}
    [Write Abstract]
\end{abstract}

\section{Introduction}
\section{Theory}
    \subsection{Normal Science}
        \subsubsection{Scientific Style}
        \subsubsection{Scientific Testing}
    \subsection{Distribution}
        \subsubsection{Von Neumann Discussion}
        \subsubsection{TODO}
\section{Methods}
    \subsection{Formula Adjustments}
        \subsubsection{Temperature Probability Adjustment}
        \subsubsection{Temperature Calculation Adjustment}
        \subsubsection{Temperature Usage Adjustment}
    \subsection{Chi \^ 2 Distribution Testing}
\section{Results}
    \subsection{Chi \^ 2 Table}
\section{Discussion}
    \subsection{Distributed Computation Accuracy}
    \subsection{Prediction}

\section{Body: Distributed Decision Making and Normal Science}

\subsection{Distributed Decision Making}

    The distributed nature of decision making is essential to modeling intelligent processes [..]

\subsection{Normal Science}

\subsection{Notes}

@@

    Even though copycat uses simulated parallel code, if copycat were actually parallelized, the global variable of temperature would actually prevent most copycat codelets from running at the same time.
    If this global variable and other constricting centralized structures were removed, copycat's code would more closely replicate intelligent processes and would be able to be run much faster.
    From a function-programming like perspective (i.e. LISP, the original language of copycat), the brain should simply be carrying out the same function in many locations (i.e. mapping neuron.process() across each of its neurons, if you will...)
    However, in violating this model with the introduction of global variables......

@@

    \break
    .....
    \break

@@

\subsubsection{Functional Programming Languages and the Brain}

@@

\subsection{Initial Formula Adjustments}

@@

After some experimentation and reading the original copycat documentation, it was clear that $S$ should be chosen to be $0.5$ and that $U$ should implement the probability curving desired at high temperatures. 
The following formulas let $U = p^r$ if $p < 0.5$ and let $U = p^\frac{1}{r}$ if $p >= 0.5$.
This controls whether/when curving happens.
Now, the parameter $r$ simply controls the degree to which curving happens.
Different values of $r$ were experimented with (values between $10$ and $1$ were experimented with at increasingly smaller step sizes. 
$2$ and $1.05$ are both good choices at  opposite "extremes".
$2$ works because it is large enough to produce novel changes in behavior at extreme temperatures without totally disregarding the original probabilities.
Values above $2$ do not work because they make probabilities too uniform.
Values below $2$ (and above $1.05$) are feasible, but produce less curving and therefore less unique behavior.
$1.05$ works because it very closely replicates the original copycat formulas, providing a very smooth curving.
Values beneath $1.05$ essentially leave probabilities unaffected, producing no significant unique behavior dependent on temperature.

@@

\subsection{Steps/plan}

Normal Science:
\begin{enumerate}
	\item Introduce statistical techniques
    \item Reduce magic number usage, document reasoning and math
    \item Propose effective human subject comparison
\end{enumerate}
Temperature:
\begin{enumerate}
	\item Propose formula improvements
    \item Experiment with a destructive removal of temperature
    \item Experiment with a "surgical" removal of temperature
    \item Assess different copycat versions with/without temperature
\end{enumerate}

\subsection{Semi-structured Notes}

\subsection{Random Notes}

This is all just free-flow unstructured notes. Don't take anything too seriously :).

Below are a list of relevant primary and secondary sources I am reviewing:

Biological/Psychological Plausibility:
\begin{verbatim}
http://www.cell.com/trends/cognitive-sciences/abstract/S1364-6613(16)30217-0
"There is no evidence for a single site of working memory storage."
https://ekmillerlab.mit.edu/2017/01/10/the-distributed-nature-of-working-memory/

Creativity as a distributed process (SECONDARY: Review primaries)
https://blogs.scientificamerican.com/beautiful-minds/the-real-neuroscience-of-creativity/
cognition results from the dynamic interactions of distributed brain areas operating in large-scale networks
http://scottbarrykaufman.com/wp-content/uploads/2013/08/Bressler_Large-Scale_Brain_10.pdf

\end{verbatim}

>>>>

\begin{abstract}
    We investigate the distributed nature of computation in a FARG architecture, Copycat. 
    One of the foundations of those models is the \emph{Parallel Terraced Scan}--a psychologically-plausible model that enables a system to fluidly move between different modes of processing.  
    Previous work has modeled decision-making under Parallel Terraced Scan by using a central variable of \emph{Temperature}. 
    However, it is unlikely that this design decision accurately replicates the processes in the human brain.
    This paper proposes several changes to copycat architectures that will increase their modeling accuracy.
\end{abstract}

\section{Introduction}

    This paper stems from Mitchell's (1993) and Hofstadter's \& FARG's (1995) work on the copycat program. 
    This project focuses on effectively simulating intelligent processes through increasingly distributed decision-making.
    In the process of evaluating the distributed nature of copycat, this paper also proposes a "Normal Science" framework.

    First, copycat uses a "Parallel Terraced Scan" as a humanistic inspired search algorithm.
    The Parallel Terraced Scan corresponds to the psychologically-plausible behavior of briefly browsing, say, a book, and delving deeper whenever something sparks one's interest. 
    In a way, it is a mix between a depth-first and breadth-first search.
    This type of behavior seems to very fluidly change the intensity of an activity based on local, contextual cues. 
    Previous FARG models use centralized structures, like the global temperature value, to control the behavior of the Parallel Terraced Scan.
    This paper explores how to maintain the same behavior while distributing decision-making throughout the system.

    Specifically, this paper attempts different refactors of the copycat architecture.
    First, the probability adjustment formulas based on temperature are changed.
    Then, we experiment with two methods for replacing temperature with a distributed metric.
    Initially, temperature is removed destructively, essentially removing any lines of code that mention it, simply to see what effect it has.
    Then, a surgical removal of temperature is attempted, leaving in tact affected structures or replacing them by effective distributed mechanisms.

    To evaluate the distributed nature of copycat, this paper focuses on the creation of a `normal science' framework.
    By `Normal science,' this paper means the term created by Thomas Kuhn--the collaborative enterprise of furthering understanding within a paradigm. 
    Today, "normal science" is simply not done on FARG architectures (and on most computational cognitive architectures too... see Addyman \& French 2012). 
    Unlike mathematical theories or experiments, which can be replicated by following the materials and methods, computational models generally have dozens of particularly tuned variables, undocumented procedures, multiple assumptions about the users computational environment, etc.
    It then becomes close to impossible to reproduce a result, or to test some new idea scientifically. 
    This paper focuses on the introduction of statistical techniques, reduction of "magic numbers", improvement and documentation of formulas, and proposals for statistical human comparison.

    We also discuss, in general, the nature of the brain as a distributed system.
    While the removal of a single global variable may initially seem trivial, one must realize that copycat and other cognitive architectures have many central structures.
    This paper explores the justification of these central structures in general.
    Is it possible to model intelligence with them, or are they harmful?

\section{Theory}

@@

    Even though copycat uses simulated parallel code, if copycat were actually parallelized, the global variable of temperature would actually prevent most copycat codelets from running at the same time.
    If this global variable and other constricting centralized structures were removed, copycat's code would more closely replicate intelligent processes and would be able to be run much faster.
    From a function-programming like perspective (i.e. LISP, the original language of copycat), the brain should simply be carrying out the same function in many locations (i.e. mapping neuron.process() across each of its neurons, if you will...)
    Note that this is more similar to the behavior of a GPU than a CPU....?
    However, in violating this model with the introduction of global variables......

@@

    \break

@@

    \break

@@

    \subsection{Normal Science}
        \subsubsection{Scientific Style}
        \subsubsection{Scientific Testing}
    \subsection{Distribution}
        \subsubsection{Von Neumann Discussion}

        \subsubsection{Turing Completeness}
        \subsubsection{Computers Can Simulate Brains}
        \subsubsection{Simulation of Distributed Processes}
        \subsubsection{Efficiency of True Distribution}
        \subsubsection{Temperature in Copycat}
        \subsubsection{Other Centralizers in Copycat}
        \subsubsection{The Motivation for Removing Centralizers in Copycat}
\section{Methods}
    \subsection{Formula Adjustments}
        \subsubsection{Temperature Probability Adjustment}

@@

            After some experimentation and reading the original copycat documentation, it was clear that $S$ should be chosen to be $0.5$ and that $U$ should implement the probability curving desired at high temperatures. 
            The following formulas let $U = p^r$ if $p < 0.5$ and let $U = p^\frac{1}{r}$ if $p >= 0.5$.
            This controls whether/when curving happens.
            Now, the parameter $r$ simply controls the degree to which curving happens.
            Different values of $r$ were experimented with (values between $10$ and $1$ were experimented with at increasingly smaller step sizes).
            $2$ and $1.05$ are both good choices at  opposite "extremes".
            $2$ works because it is large enough to produce novel changes in behavior at extreme temperatures without totally disregarding the original probabilities.
            Values above $2$ do not work because they make probabilities too uniform.
            Values below $2$ (and above $1.05$) are feasible, but produce less curving and therefore less unique behavior.
            $1.05$ works because it very closely replicates the original copycat formulas, providing a very smooth curving.
            Values beneath $1.05$ essentially leave probabilities unaffected, producing no significant unique behavior dependent on temperature.

@@

        \subsubsection{Temperature Calculation Adjustment}
        \subsubsection{Temperature Usage Adjustment}
    \subsection{$\chi^2$ Distribution Testing}
\section{Results}
    \subsection{$\chi^2$ Table}
\section{Discussion}
    \subsection{Distributed Computation Accuracy}
    \subsection{Prediction}
